Theories of deliberated choice

### 2023-08-23 simple.tex
Shorter notation

We now say that $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \in P^{I × \allhist}: ¬C({\allleadstog}) \implies \exists O \subseteq {\allleadstog} \suchthat O \implies ¬C$, meaning that
$\forall {\allleadstog} \notin C: \exists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset$.
\begin{example}
	$(\ar) \ileadstog \phi \lor  (\ar, \ar) \ileadstog \phi \lor …$ is not falsifiable.
	Given any $O$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in O} \card{\hist} + 1$. Pick any $\allleadstog$ such that $O \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	\label{ex:unfals}
	$(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
	Consider any $O \subseteq {\allleadstog}$ and let us prove that $S_O \cap C ≠ \emptyset$.
	By definition of a general observation, $O$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any ${\allleadstog'}$ such that $O \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	$\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
	Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
	\Cref{ex:unfals} concludes.
\end{example}
\begin{example}
	Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
	Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $O = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}
\begin{example}
	Given an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
	Given any ${\allleadstog} \notin C$, thus such that $\forall i: \ar \nileadstostg \phi$, any $O \subseteq {\allleadstog}$ only has finitely many $(i, \hist, p ≠ \phi)$ thus $S_O \cap C ≠ \emptyset$.
\end{example}

@@

Given $\histk \in \allhist$ and $\ar_l \in \allargs$, let $(\histk, …, \ar_l, ?)$ denote the set 
$\set{\histl \suchthat \histk \subset \histl \land \ar_l \in \histlend}$.
Given $\hist \in \allhist$, ${\ileadstog}$ and $\ar_k, \ar_l \in \allargs$, let $(\hist, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?)$ denote the set 
$\set{\histl \suchthat \exists \histk \suchthat \hist \subset \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend}$.
Given $\ar_k, \ar_l \in \allargs$, let 
$(…, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?)$ denote the set 
$\cup_{\hist \in \allhist} (\hist, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?) = 
\set{\histl \suchthat \exists \hist, \histk \suchthat \hist \subset \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend} = 
\set{\histl \suchthat \exists \histk \suchthat \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend}$.
Given ${\ileadstog}$, $p \in P$ and $S \subseteq \allargs$, let $S \ileadstog p$ stand for $\forall s \in S: s \ileadstog p$ and $s \in S \ileadstog p$ stand for $s \in \set{s \in S \suchthat s \ileadstog p}$.

@@

\begin{theorem}
	Given any $p \in P, \ar_k, \ar_l \in \allargs$, consider the following claim $C$: 
	\begin{equation}
		\label{eq:cl}
		\forall i \in I, \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l},
		\forall \hist \in (…, \ar_1, ?) \nileadstog p: (\hist, …, \ar_2, ?) \ileadstog p.
	\end{equation}
	Assume that ${\allleadsto} \in C$. Then 
	$\exists i \in I \suchthat [\ar_k \nileadstost p \land \ar_l \nileadstost p]$ 
	is equivalent to: 
	\begin{equation}
		\exists i \in I \suchthat \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l}:
		\exists \hist \in (…, \ar_1, ?) \nileadsto p \suchthat (\hist, …, \ar_2, ?) \ileadsto p.
	\end{equation}
\end{theorem}
\begin{proof}
	$\ar_k \nileadstost p$ iff $\exists \hist \suchthat \ar_k \in \histend \land \hist \nileadsto p$.
\end{proof}
\begin{remark}
	Idea.
	That consequence is unfalsifiable so we cannot include anything like this in $C$ (only something stronger, but of course a stronger statement could be false). Therefore, with no addition, we are doomed that exists some situations where no claim is suitable.
\end{remark}
\begin{theorem}[Informal]
	Assume that the NOT consequence holds and the claim holds.
	Prove that there is no suitable claim.
	If the claim is just the NOT cons and claim, then it is not fals.
	If the claim is weaker, it does not permit to prove that $\ar_k$ or $\ar_l$ is decisive. (BUT does it possibly prove that some arg is decisive? But it may be false.)
	If the claim is stronger, it may be false.
\end{theorem}
\begin{theorem}[Formal]
	Pick any $\ar_k ≠ \ar_l$ and $p \in P$ and assume that $\allleadsto$ fails the above consequence, that is, satisfies:
	\begin{equation}
		\label{eq:answerable}
		\forall i \in I, \exists \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l} \land
		\forall \hist \in (…, \ar_1, ?) \nileadsto p: (\hist, …, \ar_2, ?) \nileadsto p.
	\end{equation}
	Assume further that $\allleadsto$ satisfies \eqref{eq:cl}, thus:
	$\forall i \in I, \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l},
\forall \hist \in (…, \ar_1, ?) \nileadsto p: (\hist, …, \ar_2, ?) \ileadsto p$.
	Then, there is no suitable claim.
\end{theorem}
\begin{proof}
	Note that $\allleadsto$ satisfies, $\forall i \in I, \exists \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l} \land \forall \hist \in (…, \ar_1, ?): \hist \ileadsto p$; equivalently, $\ar_k \ileadstost p \lor \ar_l \ileadstost p$.

>>>>

We now say that $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \in P^{I × \allhist}: ¬C({\allleadstog}) \implies \exists O \subseteq {\allleadstog} \suchthat O \implies ¬C$, meaning that
$\forall {\allleadstog} \notin C: \exists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset$.
\begin{example}
	$(\ar) \ileadstog \phi \lor  (\ar, \ar) \ileadstog \phi \lor …$ is not falsifiable.
	Given any $O$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in O} \card{\hist} + 1$ if $\exists p \in P \suchthat (i, \hist, p) \in O$, and $l = 1$ otherwise. Pick any $\allleadstog$ such that $O \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	\label{ex:unfals}
	$(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
	Consider any $O \subseteq {\allleadstog}$ and let us prove that $S_O \cap C ≠ \emptyset$.
	By definition of a general observation, $O$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any ${\allleadstog'}$ such that $O \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	$\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
	Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
	\Cref{ex:unfals} concludes.
\end{example}
\begin{example}
	Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
	Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $O = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}
\begin{example}
	Given an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
	Given any ${\allleadstog} \notin C$, thus such that $\forall i: \ar \nileadstostg \phi$, any $O \subseteq {\allleadstog}$ only has finitely many $(i, \hist, p ≠ \phi)$ thus $S_O \cap C ≠ \emptyset$.
\end{example}

@@

Given $\histk \in \allhist$ and $\ar_l \in \allargs$, let $(\histk, …, \ar_l, ?)$ denote the set 
$\set{\histl \suchthat \histk \subset \histl \land \ar_l \in \histlend}$.
Given $\hist \in \allhist$, ${\ileadstog}$ and $\ar_k, \ar_l \in \allargs$, let $(\hist, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?)$ denote the set 
$\set{\histl \suchthat \exists \histk \suchthat \hist \subset \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend}$.
Given $\ar_k, \ar_l \in \allargs$, let 
$(…, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?)$ denote the set 
$\cup_{\hist \in \allhist} (\hist, \ar_k, ?, \nileadstog \phi, …, \ar_l, ?) = 
\set{\histl \suchthat \exists \hist, \histk \suchthat \hist \subset \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend} = 
\set{\histl \suchthat \exists \histk \suchthat \histk \subset \histl \land \ar_k \in \histkend \land \histk \nileadstog \phi \land \ar_l \in \histlend}$.
Given ${\ileadstog}$, $p \in P$, $s \in \allargs$ and $S \subseteq \allargs$, note that $s \in \ileadstoinvg(p)$ stand for $s \ileadstog p$ and $S \subseteq \ileadstoinvg(p)$ stand for $\forall s \in S: s \ileadstog p$.

@@

\begin{theorem}
	Given any $p \in P, \ar_k, \ar_l \in \allargs$, consider the following claim $C$: 
	\begin{multline}
		\label{eq:cl}
		\forall i \in I, \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l},
		\forall \hist \in (…, \ar_1, ?) \cap \nileadstoinvg(p): \\
		(\hist, …, \ar_2, ?) \subseteq \ileadstoinvg(p).
	\end{multline}
	Assume that ${\allleadsto} \in C$. Then 
	$\exists i \in I \suchthat [\ar_k \nileadstost p \land \ar_l \nileadstost p]$ 
	is equivalent to: 
	\begin{multline}
		\label{eq:cln}
		\exists i \in I \suchthat \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l}:\\
		\exists \hist \in (…, \ar_1, ?) \cap \nileadstoinv(p) \suchthat
		(\hist, …, \ar_2, ?) \subseteq \ileadstoinv(p).
	\end{multline}
\end{theorem}
\begin{proof}
	$\ar_k \nileadstost p$ iff $\exists \hist \suchthat \ar_k \in \histend \land \hist \nileadsto p$.
\end{proof}
\begin{remark}
	Idea.
	That consequence is unfalsifiable so we cannot include anything like this in $C$ (only something stronger, but of course a stronger statement could be false). Therefore, with no addition, we are doomed that exists some situations where no claim is suitable.
\end{remark}
\begin{theorem}
	The claim \eqref{eq:cl} and not \eqref{eq:cln}, thus, \eqref{eq:cl} and $\forall i, \ar_k \ileadstostg p \lor \ar_l \ileadstostg p$ is not falsifiable.
\end{theorem}
\begin{proof}

\end{proof}
\begin{theorem}[Informal]
		Assume that the NOT consequence holds and the claim holds.
	Prove that there is no suitable claim.
	If the claim is just the NOT cons and claim, then it is not fals.
	If the claim is weaker, it does not permit to prove that $\ar_k$ or $\ar_l$ is decisive. (BUT does it possibly prove that some arg is decisive? But it may be false.)
	If the claim is stronger, it may be false.
\end{theorem}
\begin{theorem}[Formal]
	Pick any $p \in P$ and $\ar_k ≠ \ar_l \in \allargs$ and assume that $\allleadsto$ fails \cref{eq:cln}, that is, satisfies:
	\begin{multline}
		\label{eq:answerable}
		\forall i \in I, \exists \set{\ar_1, \ar_2} \suchthat
		\set{\ar_1, \ar_2} = \set{\ar_k, \ar_l} \\ \land
		\forall \hist \in (…, \ar_1, ?) \cap \nileadstoinv(p): 
		(\hist, …, \ar_2, ?) \nsubseteq \nileadstoinv(p).
	\end{multline}
	Assume further that $\allleadsto$ satisfies \eqref{eq:cl}, thus:
	$\forall i \in I, \forall \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l},
\forall \hist \in (…, \ar_1, ?) \cap \nileadstoinv(p): (\hist, …, \ar_2, ?) \subseteq \ileadstoinv(p)$.
	Then, there is no suitable claim.
\end{theorem}
\begin{proof}
	Note that $\allleadsto$ satisfies, $\forall i \in I, \exists \set{\ar_1, \ar_2} \suchthat \set{\ar_1, \ar_2} = \set{\ar_k, \ar_l} \land \forall \hist \in (…, \ar_1, ?): \hist \ileadsto p$; equivalently, $\ar_k \ileadstost p \lor \ar_l \ileadstost p$.
